\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-01-23 papers/jss/article.Rnw
Add second reference where the mapreduce histogram pattern is used since we note we are aware of multiple applications.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@829 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Standardize on using the oxford/serial comma with lists of 3 or more. This seemed the far more prevalent convention in the file but there were 5 instances where it was not used.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@830 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%
% Local helpers to make this more compatible with R Journal style.
%
\newcommand{\CRANpkg}[1]{\pkg{#1}}
\RequirePackage{fancyvrb}
\RequirePackage{alltt}
\DefineVerbatimEnvironment{example}{Verbatim}{}
% Articles with many authors we should shorten to FirstAuthor, et al.
\shortcites{janus,dremel}
\author{Dirk Eddelbuettel\\Debian Project \And 
        Murray Stokely\\Google, Inc \And
        Jeroen Ooms\\UCLA}
\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}}

@@

%\paragraph*{Friends don't let friends use CSV!}
Data analysts and researchers often use character separated text formats such
as \texttt{CSV} \citep{shafranovich2005common} to export and import
data. However, anyone who has ever used \texttt{CSV} files will have noticed
that this method has many limitations: it is restricted to tabular data,
lacks type-safety, and has limited precision for numeric values.  Moreover,
ambiguities in the format itself frequently cause problems.  For example,
conventions on which characters is used as separator or decimal point vary by
country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
well-established and widely-supported format with the ability to define just
about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
for this complexity with comparatively large and verbose messages, and added
complexity at the parsing side (which are somewhat mitigated by the
availability of mature libraries and parsers). Because \texttt{XML} is text
based and has no native notion of numeric types or arrays, it usually not a
very practical format to store numeric datasets as they appear in statistical
applications.
%
A more modern, widely used format is \emph{JavaScript Object
  Notation} (\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
supports arrays and distinguishes 4 primitive types: numbers, strings,
booleans and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. Several \proglang{R} packages
implement functions to parse and generate \texttt{JSON} data from \proglang{R}
objects \citep{rjson,RJSONIO,jsonlite}.

@@

For added speed and efficiency, the \proglang{C++}, \proglang{Java},
and \proglang{Python} bindings to
Protocol Buffers are used with a compiler that translates a Protocol
Buffer schema description file (ending in \texttt{.proto}) into
language-specific classes that can be used to create, read, write and
manipulate Protocol Buffer messages.  The \proglang{R} interface, in contrast,
uses a reflection-based API that is particularly well-suited for
interactive data analysis.  
All messages in \proglang{R} have a single class
structure, but different accessor methods are created at runtime based
on the named fields of the specified message type, as described in the
next section.

@@

The \texttt{rexp.proto} schema supports all main \proglang{R} storage types holding \emph{data}.
These include \texttt{NULL}, \texttt{list} and vectors of type \texttt{logical}, 
\texttt{character}, \texttt{double}, \texttt{integer} and \texttt{complex}. In addition,
every type can contain a named set of attributes, as is the case in \proglang{R}. The \texttt{rexp.proto}
schema does not support some of the special \proglang{R} specific storage types, such as \texttt{function},
\texttt{language} or \texttt{environment}. Such objects have no native equivalent 
type in Protocol Buffers, and have little meaning outside the context of \proglang{R}.
When serializing \proglang{R} objects using \texttt{serialize\_pb}, values or attributes of
unsupported types are skipped with a warning. If the user really wishes to serialize these 
objects, they need to be converted into a supported type. For example, the  can use 
\texttt{deparse} to convert functions or language objects into strings, or \texttt{as.list}
for environments.

@@

One of the authors has used this design pattern for several large
scale studies of distributed filesystems \citep{janus}.

@@

%% DE Re-ordering so that we end on RProtoBuf
The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
extends the \proglang{R} system with the ability to create, read and write Protocol
Buffer message. \pkg{RProtoBuf} has been used extensively inside Google 
for the past three years by statisticians, analysts and software engineers.
At the time of this writing there are more than 300 active
users of \pkg{RProtoBuf} using it to read data from and otherwise interact
with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
other languages.

The \pkg{RProtoBuf} package provides users with the ability to generate,
parse and manipulate Protocol Buffer messages in \proglang{R}.  It is our hope that this
package will make Protocol Buffers more accessible to the \proglang{R} community, and
thereby makes a small contribution towards better integration between \proglang{R} and
other software systems and applications.

>>>>

%
% Local helpers to make this more compatible with R Journal style.
%
\newcommand{\CRANpkg}[1]{\pkg{#1}}
\RequirePackage{fancyvrb}
\RequirePackage{alltt}
\DefineVerbatimEnvironment{example}{Verbatim}{}
% Articles with many authors we should shorten to FirstAuthor, et al.
\shortcites{sciencecloud,janus,dremel}
\author{Dirk Eddelbuettel\\Debian Project \And 
        Murray Stokely\\Google, Inc \And
        Jeroen Ooms\\UCLA}
\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}}

@@

%\paragraph*{Friends don't let friends use CSV!}
Data analysts and researchers often use character separated text formats such
as \texttt{CSV} \citep{shafranovich2005common} to export and import
data. However, anyone who has ever used \texttt{CSV} files will have noticed
that this method has many limitations: it is restricted to tabular data,
lacks type-safety, and has limited precision for numeric values.  Moreover,
ambiguities in the format itself frequently cause problems.  For example,
conventions on which characters is used as separator or decimal point vary by
country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
well-established and widely-supported format with the ability to define just
about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
for this complexity with comparatively large and verbose messages, and added
complexity at the parsing side (which are somewhat mitigated by the
availability of mature libraries and parsers). Because \texttt{XML} is text
based and has no native notion of numeric types or arrays, it usually not a
very practical format to store numeric datasets as they appear in statistical
applications.
%
A more modern, widely used format is \emph{JavaScript Object
  Notation} (\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
supports arrays and distinguishes 4 primitive types: numbers, strings,
booleans, and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. Several \proglang{R} packages
implement functions to parse and generate \texttt{JSON} data from \proglang{R}
objects \citep{rjson,RJSONIO,jsonlite}.

@@

For added speed and efficiency, the \proglang{C++}, \proglang{Java},
and \proglang{Python} bindings to
Protocol Buffers are used with a compiler that translates a Protocol
Buffer schema description file (ending in \texttt{.proto}) into
language-specific classes that can be used to create, read, write, and
manipulate Protocol Buffer messages.  The \proglang{R} interface, in contrast,
uses a reflection-based API that is particularly well-suited for
interactive data analysis.  
All messages in \proglang{R} have a single class
structure, but different accessor methods are created at runtime based
on the named fields of the specified message type, as described in the
next section.

@@

The \texttt{rexp.proto} schema supports all main \proglang{R} storage types holding \emph{data}.
These include \texttt{NULL}, \texttt{list} and vectors of type \texttt{logical}, 
\texttt{character}, \texttt{double}, \texttt{integer}, and \texttt{complex}. In addition,
every type can contain a named set of attributes, as is the case in \proglang{R}. The \texttt{rexp.proto}
schema does not support some of the special \proglang{R} specific storage types, such as \texttt{function},
\texttt{language} or \texttt{environment}. Such objects have no native equivalent 
type in Protocol Buffers, and have little meaning outside the context of \proglang{R}.
When serializing \proglang{R} objects using \texttt{serialize\_pb}, values or attributes of
unsupported types are skipped with a warning. If the user really wishes to serialize these 
objects, they need to be converted into a supported type. For example, the  can use 
\texttt{deparse} to convert functions or language objects into strings, or \texttt{as.list}
for environments.

@@

One of the authors has used this design pattern for several large
scale studies of distributed storage systems \citep{sciencecloud,janus}.

@@

%% DE Re-ordering so that we end on RProtoBuf
The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
extends the \proglang{R} system with the ability to create, read, and write Protocol
Buffer messages. \pkg{RProtoBuf} has been used extensively inside Google 
for the past three years by statisticians, analysts, and software engineers.
At the time of this writing there are more than 300 active
users of \pkg{RProtoBuf} using it to read data from and otherwise interact
with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
other languages.

The \pkg{RProtoBuf} package provides users with the ability to generate,
parse, and manipulate Protocol Buffer messages in \proglang{R}.  It is our hope that this
package will make Protocol Buffers more accessible to the \proglang{R} community, and
thereby makes a small contribution towards better integration between \proglang{R} and
other software systems and applications.
